---
title: "Activity: Bird Species"
author: "Jared Hayes"
date: "June 7th 2023 "
output: html_notebook
---


# Set Up

```{r}
# Load `tidyverse` and `dcData` packages

library(tidyverse)
library(esquisse)


# Load the `OrdwayBirds` data set into your RStudio environment from the `dcData` package

data("OrdwayBirds", package = "dcData")


# Show a few commands used to inspect the data set a few different ways

OrdwayBirds%>%
  head()

OrdwayBirds%>%
  group_by(SpeciesName,Age)%>%
  summarise(count=n())

OrdwayBirds%>%
  select(SpeciesName,Age,Weight)%>%
  arrange(Weight)
```

# Activity

When you're finished with this activity, you will have a graph that shows what time of year various species appear at the Katherine Ordway Natural History Study Area in Inver Grove Heights, MN.

### Step 0

Before we begin, the book gives us some instructions to select a few key variables and clean up the date formatting.  Use the commands given in the book, but be sure to study the functions used so you know what is happening.  In this case, we select `SpeciesName`, `Month`, and `Day`. Then a `mutate()` command converts each variable to character, and then converts the resulting "character string" to a number.

```{r}
# Get the data table & clean up dates (see Data Computing eBook)

OrdwayBirds <-
  OrdwayBirds %>%
  select( SpeciesName, Month, Day ) %>%
  mutate( Month = as.numeric(as.character(Month)), 
          Day = as.numeric(as.character(Day)))

# Inspect resulting data
OrdwayBirds%>%
  head()
```



### Step 1




```{r}
# unique species in original data (including mis-spellings)

num_names<-n_distinct(OrdwayBirds$SpeciesName)
num_names

# unique species in the  list

data("OrdwaySpeciesNames", package = "dcData")

num_names_cleaned<-n_distinct(OrdwaySpeciesNames$SpeciesNameCleaned)
num_names_cleaned

```


### Step 2
```{r}
# inspect OrdwaySpeciesNames
OrdwaySpeciesNames%>%
  head()

# inner join

Corrected <-
  OrdwayBirds %>%
  inner_join( OrdwaySpeciesNames ) %>% 
  select( Species = SpeciesNameCleaned, Month, Day ) %>%
  na.omit()   


# inspect the data after the join
Corrected%>%
  head()

```


### Step 3

We want to identify the top 5-6 bird species sighted at the Katherine Ordway Natural History Study Area.

```{r}
# count sightings by species in descending order

Counts <- Corrected %>%
  group_by(Species) %>%
  summarize(Total_Captures = n()) %>%
  arrange(desc(Total_Captures))



# a hint in the text book recommends displaying the top 10 species to choose our threshold

head(Counts, n = 10)

# define a sightings threshold; produce a table with only birds belonging to major species

Threshold <- 50

Counts <- Corrected %>%
  group_by(Species) %>%
  summarize(Total_Captures = n()) %>%
  filter(Total_Captures > Threshold) %>%
  arrange(desc(Total_Captures))

Majors <- inner_join(Corrected, Counts, by = "Species")




```


### Step 4

We now want to produce a month-by-month count for each of the major species, and then construct a chart to tell the story of what time of year the various species appear.

```{r}
# prepare glyph-ready data table called `ByMonth`



```

Now use the ggplot template below to generate the desired graph. (Remark you can also use esquisser or mplot to create a line chart instead of the barchart)



```{r}
# ggplot commands

ByMonth %>% 
  ggplot(aes(x = Month, y = count, color = Species)) + 
  geom_line()

ByMonth %>% ggplot(aes(x=Month,y=count ,fill=Species)) +
  geom_bar(stat='identity',position='stack', width=.9) 

```

According to the graph, we can answer the questions posed:

##### 1. Which species are present year-round?  



##### 2. Which species are migratory, that is, primarily present in one or two seasons?




##### 3. What is the peak month for each major species? (bullet list)




##### 4. Which major species are seen in good numbers for at least 6 months of the year?



```{r}

```

